---
title: "Comparison of TE insertions"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  insertions_table: ""
  annotated_insertions: ""
  me_deletions: ""
  sample1: ""
  sample2: ""
  samples: ""
editor_options:
  chunk_output_type: console
output:
  html_document:
    highlight: tango
    number_sections: no
    theme: default
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{css, echo=F}
div.main-container {
  max-width: 90vw;
}
.note {
  display: block;
  padding: 9.5px;
  margin: 0 0 10px;
  font-size: 13px;
  line-height: 1.42857143;
  color: #333;
  background-color: #f5f5f5;
  border: 1px solid #ccc;
  border-radius: 4px;
}
```

```{r setup,include=F,eval=T}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = F)
library(magrittr)
library(data.table)
library(stringi)
thinSpace = "\u202F"
is.wholenumber <- function(x, tol = .Machine$double.eps^0.5)  {
  if (!is.numeric(x)) {return(F)}
  if (is.nan(x) | is.na(x)) {return(T)}
  abs(x - round(x)) < tol
  }
inline_hook <- function(x) {
  if (!is.wholenumber(x)) {
    format(x, digits = 2, nsmall = 1, big.mark = "\u202F")
  }
  else if (is.numeric(x)) {
    if (is.nan(x)) "NaN (division by zero?)"
    else format(x, nsmall = 0, big.mark = "\u202F")
  }
  else x
}
knitr::knit_hooks$set(inline = inline_hook)

vectorToText = function (v) {
  v = paste0(v, collapse = ", ")
  v %>% {if (stri_count_fixed(., ",") >= 2)
    stri_replace_last(., fixed = ', ', ', and ')
  else stri_replace_last(., fixed = ', ', ' and ')}
}

flextable::set_flextable_defaults(big.mark = thinSpace)

totalize = function (dt) {
  totalRow = dt %>%
    lapply(function (x) {
      if (!is.numeric(x)) {
        return("-")
      }
      else
        return(sum(x))
    })
  totalRow[[1]] = "Total"
  list(dt, totalRow) %>%
    rbindlist() %>%
    return()
}
```

```{r data}
```

```{r}
samples = strsplit(params$samples, ",")[[1]]
soi = c(params$sample1, params$sample2)
```

```{r libraries}
library(magrittr)
library(data.table)
library(stringi)
```

```{r}
chrs = paste0("chr", c(1:22, "X", "Y"))
class1 = c("Retroposon", "SINE", "LINE", "LTR")
class2 = c("DNA")
```

# Introduction

This report compares the transposable element (TE) insertions and deletions found in samples **`r params$sample1`** and **`r params$sample2`**, which are part of a cohort of `r length(samples)` samples analyzed with RetroInspector-NF.

A TE insertion is considered _shared_ between both samples when it was detected in both with the strict criterion (both callers). It is considered _unshared_ when it was detected in only one of them.

<div class="note">
__Note__: Insertion and deletion ids mentioned in this report are for internal use (not linked to any database) and will change if the pipeline is re-run.
</div>

# TE Insertions

## Summary table

The following table shows the number of shared and unshared TE insertions between samples `r params$sample1` and `r params$sample2`, broken down by TE class, family and subfamily.

```{r}
ins = readRDS(params$insertions_table)
indexes = soi %>% lapply(function (x) {
  which(samples == x)
}) %>% unlist()
insoiMask = ins$genotype %>% lapply(function (x) any(x[indexes] > 0)) %>% unlist()
insoi = ins[insoiMask]
insoi[, genotype := genotype %>% lapply(function (x) x[indexes])]
insoi[, SUPP_trusty := SUPP_mask_trusty %>% lapply(function (x) x[indexes]) %>% lapply(sum) %>% unlist()]
insoi[, SUPP_min3 := SUPP_mask_min3 %>% lapply(function (x) x[indexes]) %>% lapply(sum) %>% unlist()]

unshared_ins = insoi[SUPP_trusty > 0 & SUPP_min3 == 1]
shared_ins   = insoi[SUPP_trusty > 0 & SUPP_min3 == 2]

ins_unshared_counts = unshared_ins[, .N, by = .(repeat.class, repeat.subclass)]
ins_shared_counts   = shared_ins[, .N, by = .(repeat.class, repeat.subclass)]

ins_table = merge(ins_unshared_counts, ins_shared_counts,
                  by = c("repeat.class", "repeat.subclass"), all = TRUE)
ins_table[is.na(ins_table)] = 0
ins_table[, superclass := fifelse(repeat.class == "DNA", "Class II", "Class I")]
setnames(ins_table, c("N.x", "N.y"), c("Unshared", "Shared"))
setcolorder(ins_table, "superclass")

ins_table[order(superclass, repeat.class, repeat.subclass)] %>%
  totalize() %>%
  setnames(c("superclass", "repeat.class", "repeat.subclass"),
           c("Class", "Family", "Subfamily")) %>%
  flextable::flextable() %>%
  flextable::theme_booktabs() %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::merge_v(j = c("Class", "Family")) %>%
  flextable::border_inner_h(border = flextable::fp_border_default()) %>%
  flextable::fix_border_issues() %>%
  flextable::set_caption(caption = paste0("TE insertions shared and unshared between ",
                                           params$sample1, " and ", params$sample2))
```

Out of `r nrow(insoi[SUPP_trusty > 0])` unique TE insertions involving at least one of the two samples, `r nrow(shared_ins)` are shared and `r nrow(unshared_ins)` are unshared.

## Closer examination

Some unshared insertions may actually be present in both samples but not detected by both callers. Checking also calls that did not pass the strict criterion:

```{r}
comp = insoi[SUPP_trusty > 0, .N, by = .(SUPP_min3)][order(SUPP_min3)] %>%
  totalize()
copy(comp) %>%
  setnames(c("Present in N samples", "Number of insertions")) %>%
  flextable::flextable() %>%
  flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>%
  flextable::fix_border_issues()
```

## Annotation of unshared insertions

```{r}
annotatedInsertionsMin3 = readRDS(params$annotated_insertions)
annot = annotatedInsertionsMin3[seqId %chin% unshared_ins$seqId,
  .(seqnames, start, seqId, name, repeat.subclass, repeat.class,
    annot.gene_id, annot.symbol, annot.class, annot.subclass)][
  annot.class == "genes" & annot.subclass != "intergenic" & !is.na(annot.gene_id)][
  , .SD[1], by = .(seqId, annot.subclass, annot.symbol)]
```

<details><summary>Click to show table with `r nrow(annot)` rows.</summary>
<div class="note">__Note__: Annotation type is as described by R package annotatr: "cds": coding sequences, "promoter": regulatory region up to 1kb upstream, "1to5kb": regulatory region between 1 and 5kb upstream, "genes" and "introns": literal meaning.</div>
```{r}
annot[, .(
  Position = paste0(seqnames, ":", start %>% format(big.mark = thinSpace, trim = T)),
  seqId, annot.symbol, annot.gene_id, annot.subclass
)][stringi::stri_order(Position, numeric = T)] %>%
  setnames(c("Position", "INS Id", "Gene symbol", "Gene Id", "Annotation type")) %>%
  flextable::flextable() %>%
  flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>%
  flextable::merge_v(j = c("Position", "INS Id")) %>%
  flextable::border_inner_h(border = flextable::fp_border_default()) %>%
  flextable::fix_border_issues()
```
</details>

## Full list of unshared insertions

<details><summary>Click to show table with `r nrow(unshared_ins)` rows.</summary>
```{r}
unshared_ins[,
         .(
           Position = paste(seqnames, start %>% format(big.mark = thinSpace, trim = T), sep = ":"),
           seqId,
           `Length (bp)` = SVLEN,
           Type = paste0(repeat.class, " (", name, ")"),
           `Population allele frequency` = maf,
           `Present in sample` = genotype %>%
             lapply(function (x)
               x %>%
                 as.logical() %>%
                 soi[.]) %>%
             unlist()
         )][stringi::stri_order(Position, numeric = T)] %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::align_nottext_col(align = "center")
```
</details>

## Annotation of shared insertions

```{r}
annot_shared = annotatedInsertionsMin3[seqId %chin% shared_ins$seqId,
  .(seqnames, start, seqId, name, repeat.subclass, repeat.class,
    annot.gene_id, annot.symbol, annot.class, annot.subclass)][
  annot.class == "genes" & annot.subclass != "intergenic" & !is.na(annot.gene_id)][
  , .SD[1], by = .(seqId, annot.subclass, annot.symbol)]
```

<details><summary>Click to show table with `r nrow(annot_shared)` rows.</summary>
<div class="note">__Note__: Annotation type is as described by R package annotatr: "cds": coding sequences, "promoter": regulatory region up to 1kb upstream, "1to5kb": regulatory region between 1 and 5kb upstream, "genes" and "introns": literal meaning.</div>
```{r}
annot_shared[, .(
  Position = paste0(seqnames, ":", start %>% format(big.mark = thinSpace, trim = T)),
  seqId, annot.symbol, annot.gene_id, annot.subclass
)][stringi::stri_order(Position, numeric = T)] %>%
  setnames(c("Position", "INS Id", "Gene symbol", "Gene Id", "Annotation type")) %>%
  flextable::flextable() %>%
  flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>%
  flextable::merge_v(j = c("Position", "INS Id")) %>%
  flextable::border_inner_h(border = flextable::fp_border_default()) %>%
  flextable::fix_border_issues()
```
</details>

## Full list of shared insertions

<details><summary>Click to show table with `r nrow(shared_ins)` rows.</summary>
```{r}
shared_ins[,
         .(
           Position = paste(seqnames, start %>% format(big.mark = thinSpace, trim = T), sep = ":"),
           seqId,
           `Length (bp)` = SVLEN,
           Type = paste0(repeat.class, " (", name, ")"),
           `Population allele frequency` = maf,
           `Present in sample` = genotype %>%
             lapply(function (x)
               x %>%
                 as.logical() %>%
                 soi[.]) %>%
             unlist()
         )][stringi::stri_order(Position, numeric = T)] %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::align_nottext_col(align = "center")
```
</details>

# TE Deletions

## Summary table

The following table shows the number of shared and unshared TE deletions between samples `r params$sample1` and `r params$sample2`, broken down by TE class, family and subfamily.

```{r}
del = readRDS(params$me_deletions)

# Parsear SUPP_VEC_trusty y SUPP_VEC_min3 para las dos muestras de interes
del[, supp_vec_trusty_parsed := strsplit(SUPP_VEC_trusty, "")]
del[, supp_vec_min3_parsed   := strsplit(SUPP_VEC_min3, "")]

del[, SUPP_trusty_soi := supp_vec_trusty_parsed %>%
      lapply(function(x) sum(as.integer(x[indexes]))) %>% unlist()]
del[, SUPP_min3_soi := supp_vec_min3_parsed %>%
      lapply(function(x) sum(as.integer(x[indexes]))) %>% unlist()]

# Filtrar solo deleciones que afectan a alguna de las dos muestras
delsoi = del[SUPP_trusty_soi > 0]

unshared_del = delsoi[SUPP_trusty_soi > 0 & SUPP_min3_soi == 1]
shared_del   = delsoi[SUPP_trusty_soi > 0 & SUPP_min3_soi == 2]

del_unshared_counts = unshared_del[, .N, by = .(repeat.class, repeat.subclass)]
del_shared_counts   = shared_del[, .N, by = .(repeat.class, repeat.subclass)]

del_table = merge(del_unshared_counts, del_shared_counts,
                  by = c("repeat.class", "repeat.subclass"), all = TRUE)
del_table[is.na(del_table)] = 0
del_table[, superclass := fifelse(repeat.class == "DNA", "Class II", "Class I")]
setnames(del_table, c("N.x", "N.y"), c("Unshared", "Shared"))
setcolorder(del_table, "superclass")

del_table[order(superclass, repeat.class, repeat.subclass)] %>%
  totalize() %>%
  setnames(c("superclass", "repeat.class", "repeat.subclass"),
           c("Class", "Family", "Subfamily")) %>%
  flextable::flextable() %>%
  flextable::theme_booktabs() %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::merge_v(j = c("Class", "Family")) %>%
  flextable::border_inner_h(border = flextable::fp_border_default()) %>%
  flextable::fix_border_issues() %>%
  flextable::set_caption(caption = paste0("TE deletions shared and unshared between ",
                                           params$sample1, " and ", params$sample2))
```

Out of `r nrow(delsoi)` unique TE deletions involving at least one of the two samples, `r nrow(shared_del)` are shared and `r nrow(unshared_del)` are unshared.

## Full list of unshared deletions

<details><summary>Click to show table with `r nrow(unshared_del)` rows.</summary>
```{r}
unshared_del[,
         .(
           Position = paste(seqnames, start %>% format(big.mark = thinSpace, trim = T), sep = ":"),
           id,
           `Length (bp)` = svlen,
           Type = paste0(repeat.class, " (", name, ")"),
           `Population allele frequency` = maf
         )][stringi::stri_order(Position, numeric = T)] %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::align_nottext_col(align = "center")
```
</details>

## Full list of shared deletions

<details><summary>Click to show table with `r nrow(shared_del)` rows.</summary>
```{r}
shared_del[,
         .(
           Position = paste(seqnames, start %>% format(big.mark = thinSpace, trim = T), sep = ":"),
           id,
           `Length (bp)` = svlen,
           Type = paste0(repeat.class, " (", name, ")"),
           `Population allele frequency` = maf
         )][stringi::stri_order(Position, numeric = T)] %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::align_nottext_col(align = "center")
```
</details>
