--- date: "`r format(Sys.time(), '%d %B, %Y')`" params: insertions_table: "" annotated_insertions: "" sample1: "" sample2: "" samples: "" editor_options: 
  chunk_output_type: console
output: html_document: highlight: tango number_sections: no theme: default toc: yes toc_depth: 2 toc_float: collapsed: no smooth_scroll: yes --- --- title: 
"Comparison of TE insertions found on samples `r params$sample1` and `r params$sample2`" --- ```{css, echo=F} div.main-container {
  max-width: 90vw;
}
.note { display: block; padding: 9.5px; margin: 0 0 10px; font-size: 13px; line-height: 1.42857143; color: #333; background-color: #f5f5f5; border: 1px 
  solid #ccc; border-radius: 4px;
}
``` ```{r setup, include=F, eval=T} knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = F) thinSpace = "\u202F" is.wholenumber <- function(x, tol 
= .Machine$double.eps^0.5) {
  if (!is.numeric(x)) {return(F)} if (is.nan(x) | is.na(x)) {return(T)} abs(x - round(x)) < tol
}
inline_hook <- function(x) { if (!is.wholenumber(x)) { format(x, digits = 2, nsmall = 1, big.mark = "\u202F")
  } else if (is.numeric(x)) {
    if (is.nan(x)) "NaN (division by zero?)" else format(x, nsmall = 0, big.mark = "\u202F")
  } else x
}
knitr::knit_hooks$set(inline = inline_hook) vectorToText = function(v) { v = paste0(v, collapse = ", ") v %>% {if (stri_count_fixed(., ",") >= 2) 
    stri_replace_last(., fixed = ', ', ', and ')
  else stri_replace_last(., fixed = ', ', ' and ')}
}
flextable::set_flextable_defaults(big.mark = thinSpace) totalize = function(dt) { totalRow = dt %>% lapply(function(x) { if (!is.numeric(x)) return("-") 
    else return(sum(x))
  })
  totalRow[[1]] = "Total" list(dt, totalRow) %>% rbindlist() %>% return()
}
``` ```{r libraries} library(magrittr) library(data.table) library(stringi) ``` ```{r data} samples = strsplit(params$samples, ",")[[1]] soi = 
c(params$sample1, params$sample2) chrs = paste0("chr", c(1:22, "X", "Y")) class1 = c("Retroposon", "SINE", "LINE", "LTR") class2 = c("DNA") ins = 
readRDS(params$insertions_table) annotatedInsertionsMin3 = readRDS(params$annotated_insertions) indexes = soi %>% lapply(function(x) which(samples == x)) 
%>% unlist() insoiMask = ins$genotype %>% lapply(function(x) any(x[indexes] > 0)) %>% unlist() insoi = ins[insoiMask] insoi[, genotype := genotype %>% 
lapply(function(x) x[indexes])] insoi[, SUPP_trusty := SUPP_mask_trusty %>% lapply(function(x) x[indexes]) %>% lapply(sum) %>% unlist()] insoi[, SUPP_min3 
:= SUPP_mask_min3 %>% lapply(function(x) x[indexes]) %>% lapply(sum) %>% unlist()]
```
# Introduction
<div class="note"> __Note__: Insertion ids mentioned in this report are for internal use (not linked to any database) and will change if the pipeline is 
re-run. </div>
## TE insertions in common and not in common
```{r} comp = insoi[SUPP_trusty > 0, .N, by = .(SUPP_trusty)][order(SUPP_trusty)] %>% totalize() comp %>% copy() %>% setnames(c("Present in N samples", 
  "Number of insertions")) %>% flextable::flextable() %>% flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>% 
  flextable::fix_border_issues()
compD = comp$N ``` Out of `r compD[length(compD)]` unique TE insertions, `r ifelse(2 %in% comp$SUPP_trusty, comp[SUPP_trusty == 2]$N, 0)` are found in both 
samples, while `r ifelse(1 %in% comp$SUPP_trusty, comp[SUPP_trusty == 1]$N, 0)` are found only in one of them. ```{r} comp2 = insoi[SUPP_trusty > 0, .N, by 
= .(SUPP_min3)][order(SUPP_min3)] %>% totalize() copy(comp2) %>%
  setnames(c("Present in N samples", "Number of insertions")) %>% flextable::flextable() %>% flextable::theme_booktabs() %>% flextable::align(align = 
  "center", part = "all") %>% flextable::fix_border_issues()
```
# Unshared insertions
After this examination, `r comp2[SUPP_min3 == 1]$N` are reported as present in one sample only.
## Unshared insertions by class
```{r} unshared = insoi[SUPP_trusty > 0 & SUPP_min3 == 1] unshared[, .N, by = .(repeat.class, repeat.subclass)][ , superclass := fifelse(repeat.class == 
  "DNA", "Class II", "Class I")] %>% setcolorder("superclass") %>% totalize() %>% setnames(c("Class", "Family", "Subfamily", "Number of insertions")) %>% 
  .[order(Class, Family)] %>% flextable::flextable() %>% flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>% 
  flextable::merge_v(j = c("Family", "Class")) %>% flextable::border_inner_h(border = flextable::fp_border_default()) %>% flextable::fix_border_issues()
```
## Annotation of unshared insertions
```{r} annot = annotatedInsertionsMin3[ seqId %chin% unshared$seqId, .(seqnames, start, seqId, name, repeat.subclass, repeat.class, annot.gene_id, 
  annot.symbol, annot.class, annot.subclass)
][annot.class == "genes" & annot.subclass != "intergenic" & !is.na(annot.gene_id)][ , .SD[1], by = .(seqId, annot.subclass, annot.symbol)] ``` 
<details><summary>Click to show table with `r nrow(annot)` rows.</summary> ```{r} annot[, .(
  Position = paste0(seqnames, ":", start %>% format(big.mark = thinSpace, trim = T)), seqId, annot.symbol, annot.gene_id, annot.subclass 
)][stringi::stri_order(Position, numeric = T)] %>%
  setnames(c("Position", "INS Id", "Gene symbol", "Gene Id", "Annotation type")) %>% flextable::flextable() %>% flextable::theme_booktabs() %>% 
  flextable::align(align = "center", part = "all") %>% flextable::merge_v(j = c("Position", "INS Id")) %>% flextable::border_inner_h(border = 
  flextable::fp_border_default()) %>% flextable::fix_border_issues()
``` </details>
## Full list
<details><summary>Click to show table with `r nrow(unshared)` rows.</summary> ```{r} unshared[, .( Position = paste(seqnames, start %>% format(big.mark = 
  thinSpace, trim = T), sep = ":"), seqId, `Length (bp)` = SVLEN, Type = paste0(repeat.class, " (", name, ")"), `Population allele frequency` = maf, 
  `Present in sample` = genotype %>% lapply(function(x) x %>% as.logical() %>% soi[.]) %>% unlist()
)][stringi::stri_order(Position, numeric = T)] %>% flextable::flextable() %>% flextable::autofit() %>% flextable::theme_booktabs() %>% 
  flextable::align_nottext_col(align = "center")
``` </details>
# Shared insertions
`r ifelse(2 %in% comp2$SUPP_min3, comp2[SUPP_min3 == 2]$N, 0)` insertions are reported as shared between both samples.
## Shared insertions by class
```{r} shared = insoi[SUPP_trusty > 0 & SUPP_min3 == 2] shared[, .N, by = .(repeat.class, repeat.subclass)][ , superclass := fifelse(repeat.class == "DNA", 
  "Class II", "Class I")] %>% setcolorder("superclass") %>% totalize() %>% setnames(c("Class", "Family", "Subfamily", "Number of insertions")) %>% 
  .[order(Class, Family)] %>% flextable::flextable() %>% flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>% 
  flextable::merge_v(j = c("Family", "Class")) %>% flextable::border_inner_h(border = flextable::fp_border_default()) %>% flextable::fix_border_issues()
```
## Annotation of shared insertions
```{r} annot = annotatedInsertionsMin3[ seqId %chin% shared$seqId, .(seqnames, start, seqId, name, repeat.subclass, repeat.class, annot.gene_id, 
  annot.symbol, annot.class, annot.subclass)
][annot.class == "genes" & annot.subclass != "intergenic" & !is.na(annot.gene_id)][ , .SD[1], by = .(seqId, annot.subclass, annot.symbol)] ``` 
<details><summary>Click to show table with `r nrow(annot)` rows.</summary> ```{r} annot[, .(
  Position = paste0(seqnames, ":", start %>% format(big.mark = thinSpace, trim = T)), seqId, annot.symbol, annot.gene_id, annot.subclass 
)][stringi::stri_order(Position, numeric = T)] %>%
  setnames(c("Position", "INS Id", "Gene symbol", "Gene Id", "Annotation type")) %>% flextable::flextable() %>% flextable::theme_booktabs() %>% 
  flextable::align(align = "center", part = "all") %>% flextable::merge_v(j = c("Position", "INS Id")) %>% flextable::border_inner_h(border = 
  flextable::fp_border_default()) %>% flextable::fix_border_issues()
``` </details>
## Full list
<details><summary>Click to show table with `r nrow(shared)` rows.</summary> ```{r} shared[, .( Position = paste(seqnames, start %>% format(big.mark = 
  thinSpace, trim = T), sep = ":"), seqId, `Length (bp)` = SVLEN, Type = paste0(repeat.class, " (", name, ")"), `Population allele frequency` = maf, 
  `Present in sample` = genotype %>% lapply(function(x) x %>% as.logical() %>% soi[.]) %>% unlist()
)][stringi::stri_order(Position, numeric = T)] %>% flextable::flextable() %>% flextable::autofit() %>% flextable::theme_booktabs() %>% 
  flextable::align_nottext_col(align = "center")
``` </details>
