--- title: "TE analysis report" date: "`r format(Sys.time(), '%d %B, %Y')`" params: insertions_table: "" all_ins: "" annotated_insertions: "" me_deletions: 
  "" egoMF: "" egoBP: "" egoCC: "" do_rds: "" ncg: "" plimit: 0.05
editor_options: chunk_output_type: console output: html_document: highlight: tango number_sections: yes theme: default toc: yes toc_depth: 2 toc_float: 
      collapsed: no smooth_scroll: yes
--- ```{css, echo=F} div.main-container { max-width: 90vw;
}
.note { display: block; padding: 9.5px; margin: 0 0 10px; font-size: 13px; line-height: 1.42857143; color: #333; background-color: #f5f5f5; border: 1px 
  solid #ccc; border-radius: 4px;
}
``` ```{r setup,include=F,eval=T} knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = F) thinSpace = "\u202F" is.wholenumber <- function(x, tol = 
.Machine$double.eps^0.5) {
  if (!is.numeric(x)) {return(F)} if (is.nan(x) | is.na(x)) {return(T)} abs(x - round(x)) < tol
  }
inline_hook <- function(x) { if (!is.wholenumber(x)) { format(x, digits = 2, nsmall = 1, big.mark = "\u202F")
  }
  else if (is.numeric(x)) { if (is.nan(x)) "NaN (division by zero?)" else format(x, nsmall = 0, big.mark = "\u202F")
  }
  else x
}
knitr::knit_hooks$set(inline = inline_hook) vectorToText = function (v) { v = paste0(v, collapse = ", ") v %>% {if (stri_count_fixed(., ",") >= 2) 
    stri_replace_last(., fixed = ', ', ', and ')
  else stri_replace_last(., fixed = ', ', ' and ')}
}
totalize = function (dt) { totalRow = dt %>% lapply(function (x) { if (!is.numeric(x)) { return("-")
      }
      else return(sum(x))
    })
  totalRow[[1]] = "Total" list(dt, totalRow) %>% rbindlist() %>% return()
}
flextable::set_flextable_defaults(big.mark = thinSpace) ``` ```{r libraries} library(magrittr) library(data.table) library(stringi) library(ggplot2) 
library(patchwork) library(qqman) library(enrichplot) ``` ```{r cutomFacetWrapDefinition} scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) { stop("which must be an integer of length 1")
  }
  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) { stop("scale must be an x or y position scale")
  }
  structure(list(which = which, scale = scale), class = "scale_override")
}
CustomFacetWrap <- ggproto( "CustomFacetWrap", FacetWrap, init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) { scales <- 
    ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params) if(is.null(params$scale_overrides)) return(scales) max_scale_x <- 
    length(scales$x) max_scale_y <- length(scales$y) for(scale_override in params$scale_overrides) {
      which <- scale_override$which scale <- scale_override$scale if("x" %in% scale$aesthetics) { if(!is.null(scales$x)) { if(which < 0 || which > 
          max_scale_x) stop("Invalid index of x scale: ", which) scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) { if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which) scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }
    scales
  }
) facet_wrap_custom <- function(..., scale_overrides = NULL) { facet_super <- facet_wrap(...) if(inherits(scale_overrides, "scale_override")) { 
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) ||
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) { stop("scale_overrides must be a scale_override object or a 
    list of scale_override objects")
  }
  facet_super$params$scale_overrides <- scale_overrides ggproto(NULL, CustomFacetWrap, shrink = facet_super$shrink, params = facet_super$params )
}
chrsLength = c(248956422, 242193529, 198295559, 190214555, 181538259, 170805979, 159345973, 145138636, 138394717, 133797422, 135086622, 133275309, 
               114364328, 107043718, 101991189, 90338345, 83257441, 80373285, 58617616, 64444167, 46709983, 50818468, 156040895, 57227415)
chrScales = lapply(1:length(chrsLength), function (i) { scale_override(i, scale_x_continuous(limits = c(1, chrsLength[i])))
                     })
``` ```{r data} allIns = readRDS(params$all_ins) insertionsTable = readRDS(params$insertions_table) annotatedInsertionsMin3 = 
readRDS(params$annotated_insertions) chrs = paste0("chr", c(1:22, "X", "Y")) class1 = c("Retroposon", "SINE", "LINE", "LTR") class2 = c("DNA") ``` <div 
class="note"> __Note__: Insertion ids mentioned in this report are for internal use (not linked to any database) and will change if the pipeline is re-run. 
</div>
# General figures
## Filtering criteria
### Lax criterion
<div class="note"> __Note__: A "unique" insertion is the merged result after the allele-aware merging from both callers for each patient, and then merged 
across patients. If we were working with SNPs, a "unique" SNP would be, rs123, for example, defined by its position and alternative(s) allele(s). A unique 
insertion is defined by its position, length, and inserted sequence. </div> The analysis has identified: - `r allIns[repeat.class %in% c(class1, class2) & 
repeat.percentage >= 0.85] %>% nrow()` unique TE insertions, which amount to a total of `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 
0.85]$SUPP_min3 %>% sum()` occurrences among all samples. - With any insertion appearing on `r allIns[repeat.class %in% c(class1, class2) & 
repeat.percentage >= 0.85]$SUPP_min3 %>% mean()` ± `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85]$SUPP_min3 %>% sd()` samples 
(median of `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85]$SUPP_min3 %>% median()` occurrences/insertion). - The average 
insertion count per case is `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85]$SUPP_VEC_min3 %>% strsplit(split = "") %>% 
lapply(as.numeric) %>% simplify2array() %>% t() %>% apply(., 2, sum) %>% mean()` ± `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 
0.85]$SUPP_VEC_min3 %>% strsplit(split = "") %>% lapply(as.numeric) %>% simplify2array() %>% t() %>% apply(., 2, sum) %>% sd()`.
### Strict criterion
The stringent criterion (both callers) has left: - `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 0] %>% nrow()` 
TE insertions. - A total of `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 0]$SUPP_trusty %>% sum()` calls across 
cases. - A mean prevalence of `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 0]$SUPP_trusty %>% mean()` ± `r 
allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 0]$SUPP_trusty %>% sd()`. - The median prevalence is `r 
allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 0]$SUPP_trusty %>% median()`. - Each sample carries an average of `r 
allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 0]$SUPP_VEC_trusty %>% strsplit(split = "") %>% lapply(as.numeric) 
%>% simplify2array() %>% t() %>% apply(., 2, sum) %>% mean()` ± `r allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & SUPP_trusty > 
0][repeat.percentage >= 0.85 & SUPP_trusty > 0]$SUPP_VEC_trusty %>% strsplit(split = "") %>% lapply(as.numeric) %>% simplify2array() %>% t() %>% apply(., 2, 
sum) %>% sd()` TE insertions.
## Genotyping
```{r genotypeVsTrusty, include = F} hasAlleleleMask = insertionsTable$genotype %>% lapply(`>=`, y = 1) strictWithoutAllelesMask = mapply(function (x, y) 
{!x & y},
                                  hasAlleleleMask, insertionsTable$SUPP_mask_trusty, SIMPLIFY = F) strictWithoutAlleles = strictWithoutAllelesMask %>% 
  lapply(as.numeric) %>% lapply(sum)
strictWithFalsePositivesMask = strictWithoutAlleles >= 1 strictWithFalsePositives = insertionsTable[strictWithFalsePositivesMask] w = 
(strictWithoutAlleles[strictWithFalsePositivesMask] %>% unlist() == strictWithFalsePositives$SUPP_trusty) %>% as.numeric() %>% sum() W = 
strictWithFalsePositives[strictWithoutAlleles[strictWithFalsePositivesMask] %>% unlist() == strictWithFalsePositives$SUPP_trusty]$SUPP_trusty %>% sum() 
w_tick =
  (strictWithoutAlleles[strictWithFalsePositivesMask] %>% unlist() == strictWithFalsePositives$SUPP_trusty & strictWithFalsePositives$SUPP_trusty == 1) %>% 
  as.numeric() %>% sum()
genotypeBenchmarkTable = data.table( criterion = "Strict", falsePos = nrow(strictWithFalsePositives), falsePosPercent = nrow(strictWithFalsePositives) / 
  nrow(insertionsTable[SUPP_trusty >= 1]) * 100, occurrences = sum(unlist(strictWithoutAlleles[strictWithFalsePositivesMask])), occurrencesPercent = 
  sum(unlist(strictWithoutAlleles[strictWithFalsePositivesMask])) / sum(insertionsTable$SUPP_trusty) * 100, all0.0 = w, all0.0Occur = W)
``` - `r nrow(strictWithFalsePositives)` unique insertions that passed the stringent filtering (`r nrow(strictWithFalsePositives) / 
nrow(insertionsTable[SUPP_trusty >= 1]) * 100`% of the total unique insertions), had at least one occurrence genotyped as 0/0. - These amount to `r 
sum(unlist(strictWithoutAlleles[strictWithFalsePositivesMask]))` total count (`r sum(unlist(strictWithoutAlleles[strictWithFalsePositivesMask])) / 
sum(insertionsTable$SUPP_trusty) * 100`% of filter-passing occurrences). - `r w` of these unique insertions were genotyped as 0/0 for all their occurrences, 
amounting to `r W` insertion counts. - Of these fully wrong calls, `r w_tick` were only found on a single patient each (`r w_tick / w * 100`% of the `r w`). 
```{r genotypeVsMin3, echo = F, include = F} allelesWithoutStrictMask =
  mapply(function (x, y, z) {z >= 1 & x & !y}, hasAlleleleMask, insertionsTable$SUPP_mask_trusty, insertionsTable$SUPP_trusty, SIMPLIFY = F) 
allelesWithoutStrict = allelesWithoutStrictMask %>%
  lapply(as.numeric) %>% lapply(sum) strictWithFalseNegativesMask = allelesWithoutStrict >= 1 strictWithFalseNegatives = 
insertionsTable[strictWithFalseNegativesMask] w = (allelesWithoutStrict[strictWithFalseNegativesMask] %>% unlist() >= strictWithFalseNegatives$SUPP_trusty) 
%>% as.numeric() %>% sum() genotypeBenchmarkTable[
  , `:=`( falseNeg = nrow(strictWithFalseNegatives), falseNegPercent = nrow(strictWithFalseNegatives) / nrow(insertionsTable[SUPP_trusty >= 1]) * 100, 
    falseNegOcc = sum(unlist(allelesWithoutStrict[strictWithFalseNegativesMask])), callsThatPassed = sum(insertionsTable$SUPP_trusty), moreWrongThanRight = 
    w
  ) ] ``` - The strict criteria missed calls for `r nrow(strictWithFalseNegatives)` unique insertions (`r nrow(strictWithFalseNegatives) / 
nrow(insertionsTable[SUPP_trusty >= 1]) * 100`%). - These mean a total of `r sum(unlist(allelesWithoutStrict[strictWithFalseNegativesMask]))` occurrences, 
out of `r sum(insertionsTable$SUPP_trusty)` calls that passed the stringent filter. - `r w` unique insertions had more false negative occurrences than 
positive ones. ```{r table4, echo = F}
  min3WithoutAllelesMask = mapply(function (x, y) {!x & y}, hasAlleleleMask, insertionsTable$SUPP_mask_min3, SIMPLIFY = F) min3WithoutAlleles = 
  min3WithoutAllelesMask %>%
     lapply(as.numeric) %>% lapply(sum) laxWithFalsePositivesMask = min3WithoutAlleles >= 1 laxWithFalsePositives = 
  insertionsTable[laxWithFalsePositivesMask] w = (min3WithoutAlleles[laxWithFalsePositivesMask] %>% unlist() == laxWithFalsePositives$SUPP_min3) %>% 
  as.numeric() %>% sum() W = laxWithFalsePositives[min3WithoutAlleles[laxWithFalsePositivesMask] %>% unlist() == laxWithFalsePositives$SUPP_min3]$SUPP_min3 
  %>% sum() w_tick =
    (min3WithoutAlleles[laxWithFalsePositivesMask] %>% unlist() == laxWithFalsePositives$SUPP_min3 & laxWithFalsePositives$SUPP_min3 == 1) %>% as.numeric() 
     %>% sum()
  genotypeBenchmarkLax = data.table( criterion = "Lax", falsePos = nrow(laxWithFalsePositives), falsePosPercent = nrow(laxWithFalsePositives) / 
    nrow(insertionsTable[SUPP_min3 >= 1]) * 100, occurrences = sum(unlist(min3WithoutAlleles[laxWithFalsePositivesMask])), occurrencesPercent = 
    sum(unlist(min3WithoutAlleles[laxWithFalsePositivesMask])) / sum(insertionsTable$SUPP_min3) * 100, all0.0 = w, all0.0Occur = W)
  allelesWithoutMin3Mask = mapply(function (x, y, z) {z >= 1 & x & !y}, hasAlleleleMask, insertionsTable$SUPP_mask_min3, insertionsTable$SUPP_min3, SIMPLIFY 
           = F)
  allelesWithoutMin3 = allelesWithoutMin3Mask %>% lapply(as.numeric) %>% lapply(sum) laxWithFalseNegativesMask = allelesWithoutMin3 >= 1 
  laxWithFalseNegatives = insertionsTable[laxWithFalseNegativesMask] w = (allelesWithoutMin3[laxWithFalseNegativesMask] %>% unlist() >= 
  laxWithFalseNegatives$SUPP_min3) %>% as.numeric() %>% sum() genotypeBenchmarkLax[ , `:=`(
    falseNeg = nrow(laxWithFalseNegatives), falseNegPercent = nrow(laxWithFalseNegatives) / nrow(insertionsTable[SUPP_min3 >= 1]) * 100, falseNegOcc = 
    sum(unlist(allelesWithoutMin3[strictWithFalseNegativesMask])), callsThatPassed = sum(insertionsTable$SUPP_min3), moreWrongThanRight = w
  )] genotypeBenchmarkTable = rbind(genotypeBenchmarkTable, genotypeBenchmarkLax)[, .( Criterion = criterion, `Unique INS with false positives (%)` = 
      paste0(format(falsePos, big.mark = "\u202F"), "\n(", format(falsePosPercent, digits = 2, nsmall = 1), ")"), `Total false positives (%)` = 
      paste0(format(occurrences, big.mark = "\u202F"), "\n(", format(occurrencesPercent, digits = 1, nsmall = 1) %>% stri_trim_left() %>% unlist(), ")"), 
      `Unique INS (total) with all false positives` = paste0(all0.0 %>% format(big.mark = "\u202F"), "\n(", all0.0Occur, ")"), `Unique INS with false 
      negatives (%)` = paste0(format(falseNeg, big.mark = "\u202F"), "\n(", format(falseNegPercent, digits = 2, nsmall = 1) %>% stri_trim_left() %>% 
      unlist(), ")"), `Total false negatives (Total positives)` = paste0(falseNegOcc %>% format(big.mark = "\u202F"), "\n(", callsThatPassed %>% 
      format(big.mark = "\u202F"), ")"), `Unique INS with more false negatives than positives` = moreWrongThanRight
    ) ] genotypeBenchmarkTable %>% flextable::flextable() %>% flextable::theme_booktabs() %>% flextable::align(align = "center", part = "all") %>% 
  flextable::border_inner_h(border = flextable::fp_border_default()) %>% flextable::fix_border_issues()
``` ```{r} allFalsePositivesIds = strictWithFalsePositives[strictWithoutAlleles[strictWithFalsePositivesMask] %>% unlist() == 
  strictWithFalsePositives$SUPP_trusty]$seqId
allIns = allIns[repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & !seqId %in% allFalsePositivesIds] insertionsTable = insertionsTable[!seqId 
%in% allFalsePositivesIds] annotatedInsertionsMin3 = annotatedInsertionsMin3[!seqId %in% allFalsePositivesIds] ```
### Allele frequency
```{r results="asis"} ubiquitous = insertionsTable[maf == 1] if (nrow(ubiquitous) != 0) { cat("\n\n### Ubiquitous TE insertions\n\n") 
  cat(paste0(nrow(ubiquitous), " TE insertions were found on all cases in homozygosis.\n\n")) cat(paste0("<details><summary>Click to show table with ", 
  nrow(ubiquitous)," rows.</summary>\n\n")) ubiquitous[, .(
    Position = paste(seqnames, start %>% format(big.mark = thinSpace, trim = T), sep = ":"), Length = SVLEN, Type = paste0(repeat.class, " (", name, ")"), 
    AF = maf)] %>%
  flextable::flextable() %>% flextable::autofit() %>% flextable::theme_booktabs() %>% flextable::align_nottext_col(align = "center") %>% 
  flextable::flextable_to_rmd() cat("</details>\n\n")
}
``` - `r insertionsTable[maf >= 0.5 & SUPP_trusty > 0] %>% nrow()` TE insertions were found with an AF >= 0.5. ```{r fig.align="center"} copy = 
copy(insertionsTable[SUPP_trusty > 0])[, `:=`(seqnames = seqnames %>% as.numeric())] qqman::manhattan(copy, chr = "seqnames", bp = "start", snp = "seqId", p 
= "maf",
  logp = F, cex = 0.5, cex.axis = 0.9, ylab = "AF", chrlabs = chrs %>% sub(pattern = "chr(.+)", replacement = "\\1", x = .) %>% .[. %in% 
  as.character(unique(copy$seqnames))])
```
### Genome distribution
```{r fig.dim=c(14, 8), fig.align="center"} fig1a = insertionsTable[seqnames %in% chrs & SUPP_trusty > 0 & repeat.class %in% c(class1, class2) & 
repeat.percentage >= 0.85, .SD[1], by = c("seqId")][, seqnames := factor(seqnames, levels = chrs)] %>%
  ggplot(data = ., aes(x = start, fill = seqnames)) + geom_histogram(bins = 35) + theme_light() + theme(legend.position = "none", panel.grid.major.y = 
  element_blank(), panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = 
        element_blank(), axis.title = element_blank()) +
  facet_wrap_custom(~seqnames, ncol = 5, scales = "free_x", drop = F, scale_overrides = chrScales) fig1b = annotatedInsertionsMin3[seqnames %in% chrs & 
SUPP_trusty > 0 & repeat.class %in% c(class1, class2) & repeat.percentage >= 0.85 & annot.class == "genes" & annot.subclass %in% c("cds", "introns", 
"promoters"), .SD[1], by = c("seqId")][, seqnames := factor(seqnames, levels = chrs)] %>%
  ggplot(data = ., aes(x = start, fill = seqnames)) + geom_histogram(bins = 35) + theme_light() + theme(legend.position = "none", panel.grid.major.y = 
  element_blank(), panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = 
        element_blank(), axis.title = element_blank()) +
  facet_wrap_custom(~seqnames, ncol = 5, scales = "free_x", drop = F, scale_overrides = chrScales) fig1a + fig1b + plot_annotation(tag_levels = 'A') ```
# Analysis by TE class
```{r} roi = annotatedInsertionsMin3[repeat.class %in% c(class1, class2) & SUPP_trusty > 0 & repeat.percentage >= 0.85] ioi = roi[, .SD[1], by = seqId][, 
c("seqId", "seqnames", "start", "end", "repeat.class", "repeat.subclass", "name", "SVLEN", "SUPP_min3", "SUPP_trusty", "SUPP_VEC_trusty")] ``` ```{r} 
tableFamilies = ioi[SUPP_trusty > 0 & repeat.class %in% c("DNA", "LINE", "LTR", "Retroposon", "SINE"),
  .(Unique = .N, `Total\n(lax)` = sum(SUPP_min3), `Total\n(strict)` = sum(SUPP_trusty)), by = .(repeat.class, repeat.subclass)][order(repeat.class, 
  tolower(repeat.subclass))] %>% rbind(data.table(repeat.class = "Total", repeat.subclass = "-",
    Unique = sum(.$Unique), `Total\n(lax)` = sum(.$`Total\n(lax)`), `Total\n(strict)` = sum(.$`Total\n(strict)`))) %T>% setnames(c("repeat.class", 
  "repeat.subclass"), c("Repeat\nfamily", "Repeat\nsubfamily"))
tableFamilies %>% flextable::flextable() %>% flextable::merge_v(j = c("Repeat\nfamily")) %>% flextable::theme_booktabs() %>% 
  flextable::border_inner_h(border = flextable::fp_border_default()) %>% flextable::align(align = "center", part = "all") %>% flextable::fix_border_issues()
```
## Class I Elements
```{r fig.dim=c(14, 8), fig.align="center"} fig1a = insertionsTable[seqnames %in% chrs & SUPP_trusty > 0 & repeat.class %in% class1 & repeat.percentage >= 
0.85, .SD[1], by = c("seqId")][, seqnames := factor(seqnames, levels = chrs)] %>%
  ggplot(data = ., aes(x = start, fill = seqnames)) + geom_histogram(bins = 35) + theme_light() + theme(legend.position = "none", panel.grid.major.y = 
  element_blank(), panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = 
        element_blank(), axis.title = element_blank()) +
  facet_wrap_custom(~seqnames, ncol = 5, scales = "free_x", drop = F, scale_overrides = chrScales) fig1b = annotatedInsertionsMin3[seqnames %in% chrs & 
SUPP_trusty > 0 & repeat.class %in% class1 & repeat.percentage >= 0.85 & annot.class == "genes" & annot.subclass %in% c("cds", "introns", "promoters"), 
.SD[1], by = c("seqId")][, seqnames := factor(seqnames, levels = chrs)] %>%
  ggplot(data = ., aes(x = start, fill = seqnames)) + geom_histogram(bins = 35) + theme_light() + theme(legend.position = "none", panel.grid.major.y = 
  element_blank(), panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = 
        element_blank(), axis.title = element_blank()) +
  facet_wrap_custom(~seqnames, ncol = 5, scales = "free_x", drop = F, scale_overrides = chrScales) fig1a + fig1b + plot_annotation(tag_levels = 'A') ``` 
```{r} ioi[SUPP_trusty > 0 & repeat.subclass == "SVA", .(Name = gsub("_", " ", name), Unique = .N, `Total\n(lax)` = sum(SUPP_min3), `Total\n(strict)` = 
sum(SUPP_trusty)), by = .(name)][,-"name"][order(Name)] %>%
  totalize() %>% flextable::flextable() %>% flextable::theme_booktabs() %>% flextable::border_inner_h(border = flextable::fp_border_default()) %>% 
  flextable::align(align = "center", part = "all") %>% flextable::fix_border_issues() %>% flextable::set_caption(caption = "Count of SVA elements by name")
``` ```{r results="asis"} genic = roi[repeat.class %in% class1 & annot.subclass %in% c("cds", "introns", "promoters") & !is.na(annot.symbol), .SD[1], by = 
c("seqId", "annot.subclass")] if (nrow(genic) == 0) {
  cat("No class I insertions were found on coding sequences, introns or promoters.")
} else {
  cat(paste0("<details><summary>Click to show table with ", nrow(genic)," rows.</summary>\n\n")) genic[, .(Position = paste0(seqnames, ":", start %>% 
  format(big.mark = "\u202F", trim = T)),
    Length = SVLEN, Type = paste0(repeat.class, " (", repeat.subclass, ")"), Gene = annot.symbol, Annotation = fifelse(annot.subclass == "cds", "Coding 
    seq.", fifelse(annot.subclass == "introns", "Intron", "Promoter")), Occurrences = SUPP_trusty)][stringi::stri_order(Position, numeric = T)] %>%
  flextable::flextable() %>% flextable::autofit() %>% flextable::theme_booktabs() %>% flextable::align_nottext_col(align = "center") %>% 
  flextable::flextable_to_rmd() cat("</details>\n\n")
}
```
## Class II Elements
```{r fig.dim=c(14, 8)} fig1a = insertionsTable[seqnames %in% chrs & SUPP_trusty > 0 & repeat.class %in% class2 & repeat.percentage >= 0.85, .SD[1], by = 
c("seqId")][, seqnames := factor(seqnames, levels = chrs)] %>%
  ggplot(data = ., aes(x = start, fill = seqnames)) + geom_histogram(bins = 35) + theme_light() + theme(legend.position = "none", panel.grid.major.y = 
  element_blank(), panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = 
        element_blank(), axis.title = element_blank()) +
  facet_wrap_custom(~seqnames, ncol = 5, scales = "free_x", drop = F, scale_overrides = chrScales) fig1b = annotatedInsertionsMin3[seqnames %in% chrs & 
SUPP_trusty > 0 & repeat.class %in% class2 & repeat.percentage >= 0.85 & annot.class == "genes" & annot.subclass %in% c("cds", "introns", "promoters"), 
.SD[1], by = c("seqId")][, seqnames := factor(seqnames, levels = chrs)] %>%
  ggplot(data = ., aes(x = start, fill = seqnames)) + geom_histogram(bins = 35) + theme_light() + theme(legend.position = "none", panel.grid.major.y = 
  element_blank(), panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = 
        element_blank(), axis.title = element_blank()) +
  facet_wrap_custom(~seqnames, ncol = 5, scales = "free_x", drop = F, scale_overrides = chrScales) fig1a + fig1b + plot_annotation(tag_levels = 'A') ``` 
```{r results="asis"} genic = roi[repeat.class %in% class2 & annot.subclass %in% c("cds", "introns", "promoters") & !is.na(annot.symbol), .SD[1], by = 
c("seqId", "annot.subclass")] if (nrow(genic) == 0) {
  cat("No class II insertions were found on coding sequences, introns or promoters.")
} else {
  cat(paste0("<details><summary>Click to show table with ", nrow(genic)," rows.</summary>\n\n")) genic[, .(Position = paste0(seqnames, ":", start %>% 
  format(big.mark = "\u202F", trim = T)),
    Length = SVLEN, Type = paste0(repeat.class, " (", repeat.subclass, ")"), Gene = annot.symbol, Annotation = fifelse(annot.subclass == "cds", "Coding 
    seq.", fifelse(annot.subclass == "introns", "Intron", "Promoter")), Occurrences = SUPP_trusty)][stringi::stri_order(Position, numeric = T)] %>%
  flextable::flextable() %>% flextable::autofit() %>% flextable::theme_booktabs() %>% flextable::align_nottext_col(align = "center") %>% 
  flextable::flextable_to_rmd() cat("</details>\n\n")
}
```
# TE Deletions
```{r} meDeletionsMin3 = readRDS(params$me_deletions) ``` - `r meDeletionsMin3[SUPP_trusty > 0] %>% nrow()` TE deletions have been found, adding up to `r 
meDeletionsMin3[SUPP_trusty > 0]$SUPP_trusty %>% sum()` occurrences among cases. - These deletions have a mean prevalence per sample of `r 
meDeletionsMin3[SUPP_trusty > 0]$SUPP_trusty %>% mean()` ± `r meDeletionsMin3[SUPP_trusty > 0]$SUPP_trusty %>% sd()` (median of `r 
meDeletionsMin3[SUPP_trusty > 0]$SUPP_trusty %>% median()`). ```{r} meDeletionsMin3 = meDeletionsMin3[!(maf == 0 & SUPP_trusty > 0)] meDeletionsMin3 = 
meDeletionsMin3[stri_order(seqnames, numeric = T)] ``` ```{r fig.align="center"} copy = copy(meDeletionsMin3[SUPP_trusty > 0])[, `:=`(seqnames = 
factor(seqnames, levels = chrs) %>% as.numeric())] qqman::manhattan(copy, chr = "seqnames", bp = "start", snp = "id", p = "maf",
  logp = F, cex = 0.5, cex.axis = 0.9, ylab = "AF", chrlabs = chrs %>% sub(pattern = "chr(.+)", replacement = "\\1", x = .) %>% .[copy$seqnames %>% 
  unique()])
```
# Enrichment analysis
```{r} library(GO.db) library(clusterProfiler) library(enrichplot) library(DOSE) pLimit = params$plimit %>% as.numeric() ```
## Gene Ontology
```{r go, results="asis", fig.align="center"} fullNames = list(MF = "Molecular Function", BP = "Biological Process", CC = "Cellular Component") for (x in 
c("MF", "BP", "CC")) {
  name = paste0("ego", x) dtName = paste0(name, "Dt") paramKey = name assign(x = name, readRDS(params[[paramKey]])) object = get(name) cat("\n\n#### 
  Annotation group: ", x, " (", fullNames[[x]], ")", "\n\n", sep = "") if (is.null(object)) {
    cat("No significant results were found for this group.\n\n") next()
  }
  object@result %>% data.table() %>% .[, `:=`("geneID" = NULL, GeneRatioCalc = GeneRatio %>% lapply(function (x) str2expression(x) %>% eval()) %>% unlist() 
    )] %>% assign(dtName, ., envir = .GlobalEnv)
  dt = get(dtName)[pvalue <= pLimit] barplot(object, showCategory = 12, font.size = 10, title = paste0("GO - ", x)) %>% print() topN = min(nrow(dt), 10) 
  cat("\n\nThe category with lowest p value is *", dt[order(p.adjust)][1, Description], "* (", dt[order(p.adjust)][1, p.adjust], ")", sep = "") if (nrow(dt 
  > 1))
    cat(", and the highest (among the lowest ", topN,") is *", dt[order(p.adjust)][topN, Description] %>% as.character(), "* (", dt[order(p.adjust)][topN, 
      p.adjust],")", sep = '')
  cat('.\n\n') n = min(100, nrow(dt)) cat(paste0("<details><summary>Click to show/hide table with ", n, " rows.</summary>\n\n")) dt[order(p.adjust)][1:n] 
  %>%
    .[, lapply(.SD, function (x) if (is.numeric(x)) {sprintf("%.4g", x)} else x)] %>% flextable::flextable() %>% flextable::autofit() %>% 
  flextable::theme_booktabs() %>% flextable::align_nottext_col(align = "center") %>% flextable::flextable_to_rmd() %>% print() cat("</details>\n\n") 
  emapplot(get(name) %>% pairwise_termsim(), showCategory = 30, cex.params = list(category_label = 0.5)) %>% print() cnetplot(get(name), showCategory = 30, 
  cex.params = list(category_label = 0.5, gene_label = 0.5), node_label = "gene") %>% print()
}
```
## Disease Ontology
```{r do, results="asis"} do = readRDS(params$do_rds) if (is.null(do)) { cat("No significant results were found in DO.\n\n")
} else {
  doDt = do@result %>% data.table() if (nrow(doDt[p.adjust <= pLimit]) >= 1) { cat("Here is a barplot of most significantly enriched DO categories.\n\n") 
    barplot(do) %>% print() cat("The most significant category was '", doDt[order(p.adjust)][1, Description], "'.", sep = "")
  } else {
    cat("No significant results were found in DO.\n\n")
  }
}
```
## Network of Cancer Genes
```{r ncg, results="asis"} ncg = readRDS(params$ncg) if (is.null(ncg)) { cat("No significant results were found in NCG.\n\n")
} else {
  ncgDt = ncg@result %>% data.table() if (nrow(ncgDt[p.adjust <= pLimit]) >= 1) { cat("Here is a barplot of most significantly enriched NCG 
    categories.\n\n") barplot(ncg) %>% print() cat("The most significant category was ", ncgDt[order(p.adjust)][1, Description], ".", sep = "")
  } else {
    cat("No significant results were found in NCG.\n\n")
  }
}
```
